Finite-size scaling of free energy in the dimer model on a hexagonal domain

### 2020-01-04 AffineLieAlgebras/MQFT2018/paper.tex
Before applying Reshetikhin notes

Finite size corrections are important to close the gap between numerical simulations and
theoretical results. Moreover, they can be used to choose the proper field-theoretic description of
the scaling limit. For example, logarithmic term is supposed to be universal and dependent on the
value of central charge ${\bf c}$ of conformal field theory \cite{cardy1988finite}.

Computation of finite-size corrections to the scaling of free energy in dimer model on a rectangular
region on the square lattice led to a series of publications and even controversy about the proper CFT
description of the scaling limit, when some publications claimed ${\bf c}=-2$, while others -- ${\bf c}=1$. 
\cite{allegra2015exact,morin2016integrability,Sh_Izmailian_2019,izmailian2016finite,izmailian2005logarithmic}. 
Note that in this case limit shape does not appear due to the choice of boundary conditions.

The case of limit shape is even more complicated. While the fluctuations the limit shape can be
computed from the variational minimization problem for surface tension and fluctuations are
described by the Gaussian Free Field, the behavior on the boundary of the frozen region is not well
understood. Surface tension becomes singular near the boundary of frozen domain, thus boundary
contribution to the logarithmic correction to free energy scaling cannot be easily obtained from
field-theoretic point of view. Because of this, the computation of finite size corrections from the
exact solution of the model, especially of the logarithmic term, is very interesting.

@@

\begin{figure}[htbp]
\center{\scalebox{0.4}{\includegraphics{loz}}}
\caption{\label{dhf}A configuration of dimers on the hexagonal lattice and a corresponding picture
  of ``cubes in the corner''.}
\end{figure}

@@

Using Cardy-Peschel formulas, presented above, it should be possible to compute $\delta F$ from
value of the central charge for Gaussian free field ${\bf c}=1$, the curvature of metric, induced by
the limit shape and conformal map from the internal region of the ``Arctic curve'' to the upper
half-plane.
But the metric is singular near the boundary of the internal region and singularities are not
explicitly studied yet \cite{Astala}. Thus our result for value of $f_2$ should be useful in such a
study, which we will attempt to do in a separate publication.

>>>>

Finite-size corrections are important to fill the gap between numerical simulations and
theoretical results. Moreover, they can be used to choose the proper field-theoretic description of
the scaling limit. For example, the logarithmic term is supposed to be universal and dependent on a
value of the central charge ${\bf c}$ of conformal field theory \cite{cardy1988finite}.

Computation of finite-size corrections to the scaling of free energy in the dimer model on a rectangular
region of the square lattice led to a series of publications and even controversy about the proper CFT
description of the scaling limit. Some publications claimed ${\bf c}=-2$, while others -- ${\bf c}=1$ 
\cite{allegra2015exact,morin2016integrability,Sh_Izmailian_2019,izmailian2016finite,izmailian2005logarithmic}. 
Note that in this case the limit shape does not appear due to the choice of boundary conditions.

The case of the limit shape is even more complicated. While the limit shape can be computed from the
variational minimization problem for surface tension \cite{kenyon2007limit,kenyon2006dimers} and
fluctuations are described by the Gaussian Free Field
\cite{kenyon2001dominos,kenyon2008height,kenyon2009lectures}, the behavior on the boundary of the
frozen region is not well understood. Surface tension becomes singular near the boundary of the frozen
domain \cite{kenyon2007limit}, thus boundary contribution to the logarithmic correction to free
energy scaling cannot be easily obtained from field-theoretic point of view. As a result, the
computation of finite-size corrections from the exact solution of the model, especially of the
logarithmic term, is important.

@@

\begin{figure}[htbp]
\center{\scalebox{0.4}{\includegraphics{my-loz}}}
\caption{\label{dhf}A configuration of dimers on the hexagonal lattice and a corresponding picture
  of ``cubes in the corner''.}
\end{figure}

@@

Using Cardy-Peschel formulas, presented above, it should be possible to compute $\delta F$ from
value of the central charge for Gaussian free field ${\bf c}=1$, the curvature of metric, induced by
the limit shape and conformal map from the internal region of the ``Arctic curve'' to the upper
half-plane. But the metric is singular near the boundary of the internal region
\cite{kenyon2007limit} and the contribution of singularities to the logarithmic correction is not
explicitly studied yet. Thus our result for value of $f_2$ should be useful in such a study, which
we will attempt to do in a separate publication.
